Kung Faux Pandas \\ \large Simplifying privacy protection

### 2018-06-05 paper/main.tex
Auto Update from local Sharelatex

\usepackage{graphicx}       % added by seth - image support
\graphicspath{ {./images/} }
\usepackage{soul}           % added by seth - remove before submission, provides highlighting support
\usepackage{wrapfig}        % added by seth for placing images in-line with text

@@

The ability of independent investigators to verify results is perhaps \emph{the} most fundamental feature that separates science from other forms intellectual inquiry. Unfortunately, the mechanisms of communicating scientific results have changed little since the age of enlightenment and are inadequate for enabling reproducibility in some modern scientific endeavors such as machine learning. Perhaps the most important impediment to reproducibility is the need to share and have access to data. This has long been difficult in health care fields, but will presently become difficult for just about everyone.

There is a large body of published literature detailing methods for overcoming this difficulty related to data via de-identification (removing key identifiers, grouping individuals, and related anonymization techniques \cite{hippapro}) and synthesizing data (creating data through algorithmic means, population level statistics, randomization, and ideally preserving inter-relationships between records \cite{walonoski_synthea_2018, patki_synthetic_2016, choi_generating_2017}), however at present a mechanism to routinely \emph{use} these techniques has not been forthcoming.

This paper details an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular combination of published privacy protection methods with the popular Python Pandas data science library\cite{mckinney-proc-scipy-2010}. KFP also provides a Structured Query Language (SQL) interface which enables users to arbitrarily query any data within a database, returning to the user a data set which either does not contain any personal information though synthetic means or anonymization. We collectively refer to de-identified data and synthetic data as \emph{faux-data}. Faux in it's standard sense of artificial or fake, but more specifically as it is used in the fashion sense of the word: the use of man made materials that look and feel like real leather, fur, etc., but without the problematic derivation from live animals and reliance upon limited resources.

@@

% shorten this? Is this the right place in the background for this?
 Over the past 18 years, there has been several significant papers clarifying the definition and importance of reproducibility along with the closely related term replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result" \cite{leek_opinion_2015}; others such as Drummond use the terms in a reverse fashion \cite{drummond_replicability_2009}. Despite the semantic disagreement, both groups agree that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing details key factors in reproducibility as having: data, input parameters, documentation, software code, and an environment capable of running the provided software code; items that cannot easily nor clearly be communicated by the traditional research paper \cite{schwab_making_2000}. The previously mentioned authors propose that reproducibility should be treated as a minimum required standard that all published research should meet \cite{peng_reproducible_2006}. Specifically in the context of machine learning, this means including details to a level someone else could create the same environment including hardware configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

In the health care domain in the United States the key law governing health data is the Health Insurance Portability and Accountability Act (HIPAA). HIPAA is backed by significant civil penalties as well as specifics about security and what can and cannot be done with protected health information \cite{hippaviol}. There are two primary methods by which data can be shared or made public: de-identification, the removal of specific identifiers or summarization to a level at which individual identification is minimized, and through strict legal relationships called a business associate agreement. Even institutions which share data within the confines of the laws face a steep institutional risk. These legal barriers makes it difficult to enable independent verification of machine learning (ML) models which often require enormous amounts of training data.

Perhaps more ominously and broadly relevant for researchers, on May 25 , 2018 the European Union's (EU) General Data Protection Regulation (GDPR) has come into full effect \cite{gdpr}. This law establishes a new paradigm in data ``ownership'' for that jurisdiction. This set of rules defines the ``owners'' of data to be people from whom it was collected and obliges anyone possessing that data to honor the owners' preferences about how the data is kept and used, including a requirement to deletion of any data at the request of its owner at any time in the future.

@@

KFP addresses the issue of data access by generating faux-data that is ``statistically similar'' to the real data but does not itself contain any sensitive data. This faux-data can be analyzed, cleaned, feature-engineered, etc. as if it were the actual source data. KFP provides a standard mechanism for wrapping a privacy protection method into a plug-in which integrates the method with the Pandas DataFrame model. Two of these plug-ins are provided with full documentation for creating other plug-ins.

@@

The ideal environment for a machine learning developer to develop, test, or validate a model is to have unfettered access to all data in the problem domain they are working in. Figure~\ref{fig:architecture} depicts a basic architecture of an alternative to direct data access: data access through an intermediary that can synthesize data on-demand.

@@

Figure~\ref{fig:synthesis_process} provides a detailed look at the automated faux-data request/response process. All queries submitted to the system are first pre-processed to remove clauses such as "ORDER BY" which are made ineffective by the synthesis process. 
\begin{figure}[ht]
  \centering
  \includegraphics[width=120mm]{data_synthesis_process}
  \caption{Data synthesis process.}
  \label{fig:synthesis_process}
\end{figure}
All queries are logged for later analysis and results cached to optionally improve synthesis performance for repeated queries. Next the processed query is executed against the protected data set and results are passed into the modular synthesis process. While the current implementation only utilizes a couple of synthesis methods, machine learning developers or data warehouse owners can insert their own custom method or modify existing methods. Previously removed "ORDER BY" clauses are re-applied to the faux-data before results are returned to the user. Finally, although not implemented in this version of KFP, synthetic results could be held back until reviewed and approved for release by the data owner.

\begin{wrapfigure}[15]{r}{0.5\textwidth}%optionally place [15] after {wrapfigure} to control how many lines to wrap around figure
  \centering
  \includegraphics[width=0.5\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{wrapfigure}
KFP provides multiple methods by which a machine learning developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a machine learning process. Alternatively, an http REST service (utilized by the SPA) allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

KFP provides a unique contribution to the space of reproducible machine learning. Through providing ad-hoc faux-data, machine learning researchers can have lower barriers to accessing private data as well as reduced barriers to sharing faux-data derived from private data. However, there are several areas that could be improved or researched further: improved code sharing, addressing minimum dataset size issues, and evaluation of this system in a machine learning workflow.

While we have provided software code and a description of our methods, we believe that computational reproducibility should go beyond just making source code available. Peng recommends making code available in any form as an excellent first step towards reproducible research; the next step is to make code available "in a durable non-proprietary format" \cite{peng_reproducible_2011}. This is a level of code sharing rarely achieved in machine learning research. Currently there is no standard long term format that works for all computational environments. For the Python ecosystem however, the standard for sharing code is via a package on Python Package Index (PyPI). We intend to continue this research to make KFP available as a Python package.

Another area that is open for further research is that of privacy protection through use of minimum dataset size. In the field of education, statistical methods for calculating sample size based on probability of detecting a specific effect size are commonly used \cite{naep_2009}. However, it is not clear that these methods are appropriate for use machine learning in general or the specific domain of health care. Furthermore, effect size and probability to detect specific effects may vary by use case. Thus, in this work, no attempt has been made to constrain data set sizes or results returned. KFP could be modified to return no results if a minimum threshold is not met based on the risk profile of the data owner or, alternatively, data boosting techniques in combination with data synthesis could be used.

The last area recommended for further research is to deploy and observe the use of KFP in an active machine learning environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.

@@

Finally, instead of taking the computational and labor intensive step of full EHR de-identification or synthesis, KFP takes the smaller and more easily implemented step of just-in-time data de-identification or synthesis. We surmise that this tool will improve data exploration and reduce dependency on complex processes required for data access. The result of this system will be improved reproducibility in machine learning.

>>>>

\usepackage{graphicx}       % added by seth - image support
\graphicspath{ {./images/} }
\usepackage{wrapfig}        % added by seth for placing images in-line with text

@@

Independent reproduction and replication of results are critical components of scientific inquiry. Barriers to data access and sharing are the most important impediments to reproducibility in machine learning (ML) research. Health care research has unique challenges because of necessary personal data privacy protections.

Several methods exist to de-identify (remove key identifiers, group individuals, and related anonymization techniques \cite{hippapro}) and synthesize data (create data through algorithmic means, population level statistics, randomization, ideally with preservation of multivariable relationships between records \cite{walonoski_synthea_2018, patki_synthetic_2016, choi_generating_2017}). However, no pipeline currently exists with which ML investigators can routinely \emph{anonymize} or \emph{synthesize} data to facilitate access and sharing. This work addresses that gap.

We have created an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular combination of established privacy protection methods with the popular Python Pandas data science library\cite{mckinney-proc-scipy-2010}. KFP also provides a Structured Query Language (SQL) interface which enables users to query a database and receive a data set without personal information through either anonymization or synthetic data generation.

@@

Recent work has clarified the definition and importance of reproducibility and a closely related concept, replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result."\cite{leek_opinion_2015,peng_reproducible_2006} Others have used the terms in a reverse fashion.\cite{drummond_replicability_2009} Despite the semantic differences, there is broad agreement that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing lists the key factors in reproducibility as available data, input parameters, documentation, software code, and an environment capable of running the provided software code. These items are difficult to completely convey in a traditional research paper \cite{schwab_making_2000}.

In the context of ML, sufficient detail is required such that an independent investigator could create the same hardware and software configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

In the health care domain in the United States the key law governing health data is the Health Insurance Portability and Accountability Act (HIPAA). HIPAA is backed by significant civil penalties as well as specifics about security and what can and cannot be done with protected health information \cite{hippaviol}. There are two primary methods by which data can be shared or made public: de-identification, the removal of specific identifiers or summarization to a level at which individual identification is minimized, and through strict legal relationships called a business associate agreement. Even institutions which share data within the confines of the laws face a steep institutional risk. These legal barriers makes it difficult to enable independent verification of ML models which often require enormous amounts of training data.

Perhaps more ominously and broadly relevant for researchers, on May 25, 2018 the European Union's (EU) General Data Protection Regulation (GDPR) has come into full effect \cite{gdpr}. This law establishes a new paradigm in data ``ownership'' for that jurisdiction. This set of rules defines the ``owners'' of data to be people from whom it was collected and obliges anyone possessing that data to honor the owners' preferences about how the data is kept and used, including a requirement to deletion of any data at the request of its owner at any time in the future.

@@

KFP addresses the issue of data access by generating data that is ``statistically similar'' to the real data but does not itself contain any sensitive data. We collectively refer to de-identified data and synthetic data as \emph{faux-data}. We use the term faux as it is used in the fashion sense of the word: the use of man made materials that look and feel like real leather, fur, etc., but without the problematic derivation from live animals and reliance upon limited resources. This faux-data can be analyzed, cleaned, feature-engineered, etc. as if it were the actual source data. KFP provides a standard mechanism for wrapping a privacy protection method into a plug-in which integrates the method with the Pandas DataFrame model. Two of these plug-ins are provided with full documentation for creating other plug-ins.

@@

Ideally a ML investigator should develop, test, and validate models using all available data. This is most easily accomplished when they have unfettered access to all data in the problem domain they are working in. Figure~\ref{fig:architecture} shows the novel architecture of KFP: data access through an intermediary that can synthesize data on-demand.

@@

Figure~\ref{fig:synthesis_process} shows the automated faux-data request/response process. All queries submitted to the system are first pre-processed to remove clauses such as "ORDER BY" which are made ineffective by the synthesis process. 
\begin{figure}[ht]
  \centering
  \includegraphics[width=120mm]{data_synthesis_process}
  \caption{Data synthesis process.}
  \label{fig:synthesis_process}
\end{figure}
All queries are logged for later analysis and results cached to optionally improve synthesis performance for repeated queries. Next the processed query is executed against the protected data set and results are passed into the modular synthesis process. While the current implementation only utilizes a couple of synthesis methods, ML developers or data warehouse owners can insert their own custom method or modify existing methods. Previously removed "ORDER BY" clauses are re-applied to the faux-data before results are returned to the user. Finally, although not implemented in this version of KFP, synthetic results could be held back until reviewed and approved for release by the data owner.

\begin{wrapfigure}[15]{r}{0.5\textwidth}%optionally place [15] after {wrapfigure} to control how many lines to wrap around figure
  \centering
  \includegraphics[width=0.5\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{wrapfigure}
KFP provides multiple methods by which a ML developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a ML process. Alternatively, a Hypertext Transfer Protocol (HTTP) Representational State Transfer (REST) service \cite{w3c_working_group_webservices}, utilized by the SPA, allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

KFP provides a unique contribution to the space of reproducible ML. Through providing ad-hoc faux-data, ML researchers can have lower barriers to accessing private data as well as reduced barriers to sharing faux-data derived from private data. However, there are several areas that could be improved or researched further: improved code sharing, addressing minimum dataset size issues, and evaluation of this system in a ML workflow.

While we have provided software code and a description of our methods, we believe that ML reproducibility should go beyond just making source code available. Peng recommends making code available in any form as an excellent first step towards reproducible research; the next step is to make code available "in a durable non-proprietary format" \cite{peng_reproducible_2011}. This is a level of code sharing rarely achieved in ML research. Currently there is no standard long term format that works for all computational environments. For the Python ecosystem however, the standard for sharing code is via a package on Python Package Index (PyPI). We intend to continue this research to make KFP available as a Python package.

Another area that is open for further research is that of privacy protection through use of minimum dataset size. In the field of education, statistical methods for calculating sample size based on probability of detecting a specific effect size are commonly used \cite{naep_2009}. However, it is not clear that these methods are appropriate for use ML in general or the specific domain of health care. Furthermore, effect size and probability to detect specific effects may vary by use case. Thus, in this work, no attempt has been made to constrain data set sizes or results returned. KFP could be modified to return no results if a minimum threshold is not met based on the risk profile of the data owner or, alternatively, data boosting techniques in combination with data synthesis could be used.

The last area recommended for further research is to deploy and observe the use of KFP in an active ML environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.

@@

Finally, instead of taking the computational and labor intensive step of full EHR de-identification or synthesis, KFP takes the smaller and more easily implemented step of just-in-time data de-identification or synthesis. We surmise that this tool will improve data exploration and reduce dependency on complex processes required for data access. The result of this system will be improved reproducibility in ML.
